### 2019-05-28 doc/chapters/using_stainless.tex
Added Stainless library.

\subsection{Predefined types}

@@

\begin{itemize}
  \item \textbf{Mutable variables} which can be declared with the keyword \textit{var}.
  Declaration of mutable variables is allowed only as a local variable within a function or in a case classe constuctor.
  \item  \textbf{While loops}. 
  Because besides main calculations in a loop the state of a loop counter changes at every iteration as a side effect, this kind of programming component does not belong to the functional paradigm and cannot be handled by solvers.
  Thus, Stainless generates automatically a postcondition with the negation of the loop condition as an argument. 
  Moreover, to make it stronger, the automatically generated postcondition can be customised with the keyword \textit{invariant}.
  An example of while loop implementation:
  \begin{lstlisting}[language=Scala]
    var res = 0
    var i = 0
    (while(i < 10) {
      res = res + i
      i = i + 1
    }) invariant(i >= 0 && res >= i)
  \end{lstlisting}
  \item \textbf{Arrays.} Stainless extends functional arrays allowing the reassignment of arrays elements.
  \item ...
\end{itemize}

>>>>

@@

\begin{itemize}
  \item \textbf{Mutable variables} which can be declared with the keyword \textit{var}.
  Declaration of mutable variables is allowed only as a local variable within a function, or in a case classe constuctor, or in trait if a default value is not assigned.
  \item  \textbf{While loops}. 
  Because besides main calculations in a loop the state of a loop counter changes at every iteration as a side effect, this kind of programming component does not belong to the functional paradigm and cannot be handled by solvers.
  Thus, Stainless generates automatically a postcondition with the negation of the loop condition as an argument. 
  Moreover, to make it stronger, the automatically generated postcondition can be customised with the keyword \textit{invariant}.
  An example of while loop implementation:
  \begin{lstlisting}[language=Scala]
    var res = 0
    var i = 0
    (while(i < 10) {
      res = res + i
      i = i + 1
    }) invariant(i >= 0 && res >= i)
  \end{lstlisting}
  \item \textbf{Function \textit{old}} can be used in postconditions. 
  It allows to get value of a variable before the execution of the block.
  This feature is useful dealing with mutable objects.
  An example of using the function \textit{old}:
  \begin{lstlisting}[language=Scala]
    case class A(var x: Int)
    def inc(a: A): Unit = {
      a.x = a.x + 1
    } ensuring(_ => a.x == old(a).x + 1)
  \end{lstlisting}
  \item \textbf{Arrays.} Stainless extends functional arrays allowing the reassignment of arrays elements. \cite{Stainless:imperative}
\end{itemize}

\subsection{Stainless library}

Besides annotations and functions helping for verification  some core data types and operation on them is defined in the Stainless library.
The own implementation and dedicated mapping of data types allows Stainless to transform it correctly into mathematical expressions of SMT solvers.
Thus, the functionality of some data types can be restricted.

For example, BigInt implemented in Stainless has constructors only with a parameter of the type Int or String, while other types of a parameter is allowed in Scala.
Furthermore, Stainless defines its own collection List[T] with the standard and some additional functions because SMT solvers do not support it.
Moreover, the own Sets and Maps are used in Stainless. \cite{Stainless:library}
